### 2021-07-08 rad-icfp-2021/unf.tex
minor
reorganizing sections 5 and 6
cost model figure
more improvement correctness

There are a few notable things in this syntax. 
Every primitive is indexed by a list of types from the source language, 
which corresponds to a context being carried. 
Every constant, unary, or binary operator from Source has a corresponding $n$-ary operator !op$_{T,n}$! in UNF.
Sequential composition is denoted by !;! and !e$_{1}$;e$_{2}$! means that !e$_{1}$! should be performed, and then !e$_{2}$!.
The array operators !map2! and !reduce! have extra indices that represent well-formed terms in Source.
The language also contains a projection !proj! which forgets some of the elements, 
and a pairing operator !pair! which takes the last pair of elements and returns the pairing of these elements.

@@

\begin{example}
    The term in Source UNF !cos$_{\reals,\reals}$ $\comp$ pair$_{\reals,\reals; \reals\times\reals}$! 
    intuitively represents a term \\
    !x$_1$:$\reals$,x$_2$:$\reals$,x$_3$:$\reals$ $\vdash$ let (x$_1$,x$_2$,x$_3$,x$_4$)=(x$_1$,x$_2$,x$_3$,cos(x$_3$)) in (x$_1$,x$_2$,<x$_3$,x$_4$>)!.\\
    The operator $\comp$ can more generally be understood as a let binding of a tuple, and primitives only act on the last parts of the tuple, always also returning their whole context.
\end{example}

@@

    \begin{tabular}{r c l}
        !$\Dsynrevsymbol_{\rho}$(cos$_{\reals,\reals}\comp$ pair$_{\reals,\reals; \reals\times\reals}$)!
        &=& !<> $\comp$! \\
        &&  !<> !
    \end{tabular}
\end{example}

@@

% TODO: we could merge everything of the form op;proj into just op by changing the type of op.

@@

\begin{example}
    TODO
\end{example}

@@

\begin{example}
    TODO
\end{example}

>>>>

There are a few notable things in this syntax. 
Every primitive is indexed by a list of types from the source language, 
which corresponds to a context being carried. We will often elide the brackets ![!,!]!.
Every constant, unary, or binary operator from Source has a corresponding $n$-ary operator !op$_{T,n}$! in UNF.
Sequential composition is denoted by !;! and !e$_{1}$;e$_{2}$! means that !e$_{1}$! should be performed, and then !e$_{2}$!.
The array operators !map2! and !reduce! have extra indices that represent well-formed terms in Source.
The language also contains a projection !proj! which forgets some of the elements, 
and a pairing operator !pair! which takes the last pair of elements and returns the pairing of these elements.

@@

\begin{example}
    \label{exm:unf}
    The term in Source UNF !cos$_{\reals,\reals}$ $\comp$ pair$_{\reals,\reals; \reals\times\reals}$! 
    intuitively represents a term \\
    !x$_1$:$\reals$,x$_2$:$\reals$,x$_3$:$\reals$ $\vdash$ let (x$_1$,x$_2$,x$_3$,x$_4$)=(x$_1$,x$_2$,x$_3$,cos(x$_3$)) in (x$_1$,x$_2$,<x$_3$,x$_4$>)!.\\
    The operator $\comp$ can more generally be understood as a let binding of a tuple, and primitives only act on the last parts of the tuple, always also returning their whole context.
\end{example}

@@

    \begin{tabular}{l}
        !$\reals$,$\reals$,$\reals$,$\reals\times\reals\times\reals$->$\rho \vdash \Dsynrevsymbol_{\rho}$(cos$_{\reals,\reals}\comp$ pair$_{\reals,\reals; \reals\times\reals}$): $\reals$,$\reals$,$\reals \times \reals$,$(\reals\times\reals\times(\reals\times\reals))$->$\rho$! \\
        = !<proj$_{\reals^3;\reals^3->\rho;[]} \comp$ cos$_{\reals,\reals}$, proj$_{[];\reals^3;\reals^3->\rho} \icomp$ (proj$_{\reals^3;\reals^3->\rho;[]}$ $\comp$ J$^T$cos$_{\reals,\reals}$)> $\comp$! \\
        !<proj$_{\reals^4;\reals^4->\rho;[]} \comp$ pair$_{\reals,\reals; \reals\times\reals}$, proj$_{[];\reals^4;\reals^4->\rho} \icomp$ (proj$_{\reals^4;\reals^4->\rho;[]}$ $\comp$ J$^T$pair$_{\reals,\reals; \reals\times\reals}$)>!
    \end{tabular}
\end{example}

@@

@@

@@

\begin{example}
    Continuing again with our previous example !cos$_{\reals,\reals}$ $\comp$ pair$_{\reals,\reals; \reals\times\reals}$!, 
    we have, after removing a few unnecessary let-bindings:

    \begin{tabular}{l}
        !x$_1$,x$_2$,x$_3$:$\reals$,x$_4$:$\reals^3$->$\rho$ $\vdash$ $\UNFSymbol^{-1}$($\Dsynrevsymbol_{\rho}$(cos$_{\reals,\reals}$ $\comp$ pair$_{\reals,\reals; \reals\times\reals}$): $(\reals\times\reals) \times (\reals\times\reals\times(\reals\times\reals))$->$\rho$! \\
        = !TODO!
    \end{tabular}
\end{example}
